Elucidating the Transcriptome of Turkey Hemorrhagic Enteritis Virus

### 2023-11-09 manuscript_thev_transcriptome.Rmd
validated mlp juncs:D,F,G,H

**Major Late Promoter Region (MLP) transcripts**.

@@

### Computational Analysis of RNA Sequencing Data: Mapping and Transcript characterization
Analysis of our sequence reads were analyzed following a well established protocol described by Pertea _et al_ [@Pertea2016], using `SNAKEMAKE 7.24.0` to drive the pipeline. Briefly, sequencing reads were trimmed with the `FastQC - version 0.11.9` [@FastQC] program to achieve an overall `Mean Sequence Quality (Phred Score)` of 36. Trimmed reads were mapped to the complete sequence of avirulent turkey hemorrhagic enteritis virus strain Virginia (https://www.ncbi.nlm.nih.gov/nuccore/AY849321.1/) and *Meleagris gallopavo* (https://www.ncbi.nlm.nih.gov/genome/?term=Meleagris+gallopavo) using `Hisat2 - version 2.2.1` [@Pertea2016] with default settings without relying on known splice sites. The generated `BAM` files from each infection time-point were filtered for reads mapping to the THEV genome and fed into `StringTie - version 2.2.1` [@Pertea2016] using a `gff3` file from NCBI containing the predicted ORFs of THEV as a guide. A custom script was used to consolidate all transcripts from all time-points without redundancy, generating the final transcriptome of THEV.

### Validation of Novel Splice Junctions 
All splice junctions identified in this work are novel except one predicted splice site each for pTP and DBP, which were corroborated in our work. However, these predicted splice junctions have not been experimentally validated hitherto, and we identified additional novel splice junctions beside the predicted junctions, giving a more complete picture of the transcripts.

The novel splice junctions after consolidating all transcripts with `StringTie` which we validated by PCR and Sanger Sequencing are shown in ~~Table###1~~. We designed primers that crossed a range of novel exon–exon boundaries for each specific transcript in a transcription unit with their respective universal primers (~~supplementary, PCR methods~~). Each forward primer contained a KpnI restriction site and reverse primers, an XbaI site. After first-strand cDNA synthesis with SuperScript™ III First-Strand Synthesis System (ThermoFisher SCIENTIFIC), these primers were used in a targeted PCR experiment, the PCR products were analysed on Agarose gels, cloned by traditional restriction enzyme method and Sanger sequenced to validate these splice junctions at the sequence level.

@@

## DISCUSSION/CONCLUSIONS
In the original study where the ORFs of THEV were predicted, ORF4 was predicted in the E1 region spanning the Hyd gene. However, later studies predicted and preferred Hyd instead of ORF4; hence, the current prediction map. However, this study shows that while both Hyd and ORF4 may be both expressed, ORF4 is most likely the bona fide gene.

@@

## ACKNOWLEDGMENTS  
LC Sciences
`Eton Bioscience, Inc, San Diego, CA`
\newpage

>>>>

**Major Late Promoter Region (MLP) transcripts**.
\newpage
## DISCUSSION/CONCLUSIONS
In the original study where the ORFs of THEV were predicted, ORF4 was predicted in the E1 region spanning the Hyd gene. However, later studies predicted and preferred Hyd instead of ORF4; hence, the current prediction map. However, this study shows that while both Hyd and ORF4 may be both expressed, ORF4 is most likely the bona fide gene.
\newpage

@@

### Validation of Novel Splice Junctions 
All splice junctions identified in this work are novel except one predicted splice site each for pTP and DBP, which were corroborated in our work. However, these predicted splice junctions have not been experimentally validated hitherto, and we identified additional novel exons, giving a more complete picture of the transcripts.
~~The novel splice junctions after consolidating all transcripts with `StringTie` which we validated by PCR and Sanger Sequencing are shown in Table###1~~. We designed primers that crossed a range of novel exon–exon boundaries for each specific transcript in a transcription unit paired with their respective universal primers (~~supplementary, PCR methods~~). Each forward primer contained a KpnI restriction site and reverse primers, an XbaI site. After first-strand cDNA synthesis with SuperScript™ III First-Strand Synthesis System (ThermoFisher SCIENTIFIC), these primers were used in a targeted PCR experiment, the PCR products were analysed on Agarose gels, cloned by traditional restriction enzyme method and Sanger sequenced to validate these splice junctions at the sequence level.

@@

### Computational Analysis of RNA Sequencing Data: Mapping and Transcript characterization
Analysis of our sequence reads were analyzed following a well established protocol described by Pertea _et al_ [@Pertea2016], using `SNAKEMAKE 7.24.0` to drive the pipeline. Briefly, sequencing reads were trimmed with the `FastQC - version 0.11.9` [@FastQC] program to achieve an overall `Mean Sequence Quality (Phred Score)` of 36. Trimmed reads were mapped to the complete sequence of avirulent turkey hemorrhagic enteritis virus strain Virginia (https://www.ncbi.nlm.nih.gov/nuccore/AY849321.1/) and *Meleagris gallopavo* (https://www.ncbi.nlm.nih.gov/genome/?term=Meleagris+gallopavo) using `Hisat2 - version 2.2.1` [@Pertea2016] with default settings without relying on known splice sites. The generated `BAM` files from each infection time-point were filtered for reads mapping to the THEV genome and fed into `StringTie - version 2.2.1` [@Pertea2016] using a `gff3` file from NCBI containing the predicted ORFs of THEV as a guide. A custom script was used to consolidate all transcripts from all time-points without redundancy, generating the final transcriptome of THEV.
\newpage

@@

## ACKNOWLEDGMENTS  
`LC Sciences` - RNA sequencing was done here
`Eton Bioscience, Inc, San Diego, CA` - All Sanger sequencing validations was done here
\newpage
